---
title: "EDA"
author: "Zachariah Freitas"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.path = "README_figs/README-",
  echo = TRUE)
library(zoo)
library(tidyverse)
library(fpp2) # Plot and Forecast Data
set.seed(506)
library(readxl)
library(feasts)
library(tsibble)
# library(tsibbledata)
library(dplyr)
library(ggplot2)
library(lubridate)

df <- read_excel("Sales_Forecasting.xlsx", sheet = "Sales_Data")

# Convert Quarter character to date
df$date <- as.Date(df$svc_agreement_activation_date)
```

## EDA


```{r}

# Create Time Series Object
myts <- ts(df$activation_count, 
           start = c(2021, as.numeric(format(df$date[1], "%j"))),
           frequency = 364)


autoplot(myts) +
  xlab("Date") + 
  ylab("Activations") +
  ggtitle("Forecasts for daily activations") +
  theme_minimal() +
  guides(colour=guide_legend(title="Forecast"))



```


```{r}

temp <- df %>% 
  mutate(svc_agreement_activation_date = as.Date(svc_agreement_activation_date)) %>% 
  as_tsibble(., index = svc_agreement_activation_date)

temp %>% 
  gg_season(activation_count, period = "week") +
  theme(legend.position = "right") +
  labs(y="Activations", x="Day of Week", title="Activation Counts by Day")
```




```{r 56 weeks models}
library(forecast)

myts <- ts(df$activation_count, 
           start = c(1, 1),
           frequency = 7)

# Split Data
train <- window(myts, end = c(57, 4)) #273 orig 300
valid <- window(myts, start = c(57, 5)) # October of 2022

# Set Forecast periods
h <- length(valid)


#############
# Train Models
# Trend + Seasonal Model
ST <- forecast::forecast(forecast::tslm(train ~ trend + season),h=h)

# Exponential Model
EXP <- forecast::forecast(forecast::tslm(train ~ trend + season, lambda = 0),h=h)

# NNAR - Autofit neural network
NNAR <- forecast::forecast(nnetar(train), h=h)#, lambda=0)


# Auto ARIMA
ARIMA <- forecast::forecast(auto.arima(train, lambda=0, biasadj=TRUE),h=h)


# Ensemble Approach
Combination <- (EXP[["mean"]] + 
                  ARIMA[["mean"]] +
                  ST[["mean"]] + 
                  NNAR[["mean"]])/4

```


```{r Model 56 Weeks Plots}
############
# Baseline Naïve Model

naive_df <- df %>% 
  mutate(naive_forecast = (lag7 + lag14)/2) %>% 
  select(svc_agreement_activation_date, naive_forecast, activation_count)


# Create Time Series Object
mynaivets <- ts(naive_df$naive_forecast, 
                start = c(1,1),
                frequency = 7)

# Plot Results
autoplot(window(myts, start=c(55, 1))) +
  autolayer(window(mynaivets, start=c(57, 5)), series="Baseline Forecast") +
  autolayer(ST, series="Seasonal Trend", PI=FALSE) +
  autolayer(EXP, series="Exponential", PI=FALSE) +
  autolayer(ARIMA, series="ARIMA", PI=FALSE) +
  autolayer(NNAR, series="NNAR", PI=FALSE) +
  autolayer(Combination, series="Combination") +
  xlab("Date") + 
  ylab("Activations") +
  ggtitle("Forecasts for daily activations - 56 Weeks of Training Data") +
  theme_minimal() +
  guides(colour=guide_legend(title="Forecast"))
```

## Including Plots
```{r 56 weeks results}

# Baseline model

resids <- naive_df %>% 
  filter(svc_agreement_activation_date >= as.Date('2022-10-01')) %>% 
  mutate(residuals = naive_forecast - activation_count) %>% 
  select(residuals)

# RMSE
print("Baseline Forecast")
print("RMSE - Test Set")
round(sqrt(mean(resids$residuals^2)),2)
print("")

print(strrep("#", 80))
print("Seasonal Trend")
round(forecast::accuracy(ST, valid),2)
print("")

print(strrep("#", 80))
print("Exponential")
round(forecast::accuracy(EXP, valid),2)
print("")

print(strrep("#", 80))
print("Neural Network")
round(forecast::accuracy(NNAR, valid),2)
print("")    
    
print(strrep("#", 80))
print("ARIMA")
round(forecast::accuracy(ARIMA, valid),2)
print("")

print(strrep("#", 80))
print("Combination")
round(forecast::accuracy(Combination, valid),2)
print("")
```

You can also embed plots, for example:

```{r 56 weeks RMSE}
c(
  Baseline = round(sqrt(mean(resids$residuals^2)),2),
  ST = forecast::accuracy(ST, valid)["Test set","RMSE"],
  EXP = forecast::accuracy(EXP, valid)["Test set","RMSE"],
  NNAR = forecast::accuracy(NNAR, valid)["Test set","RMSE"],
  ARIMA = forecast::accuracy(ARIMA, valid)["Test set","RMSE"],
  Combination = forecast::accuracy(Combination, valid)["Test set","RMSE"])

```

Let's try adding the Baseline Forecast to the ensemble to see if we get an improvement in our prediction.


```{r 56 weeks combo}


baseline.predict <- df %>% 
  select(svc_agreement_activation_date, activation_count, lag7, lag14) %>% 
  mutate(naive_forecast = (lag7 + lag14)/2) %>% 
  filter(svc_agreement_activation_date >= as.Date('2022-10-01')) 


NewCombination <- (EXP[["mean"]] + 
                  ARIMA[["mean"]] +
                  ST[["mean"]] + 
                  NNAR[["mean"]] + 
                  baseline.predict[["naive_forecast"]])/5

round(forecast::accuracy(NewCombination, valid),2)

```

# Try Shorter Lookback Windows
Seeing that the better performing model uses a smaller lookback period I want to try the same models with a smaller look back period.


```{r models}
library(forecast)

myts <- ts(df$activation_count, 
           start = c(1, 1),
           frequency = 7)

# Split Data
train <- window(myts, start = c(45, 5), end = c(57, 4)) #273 orig 300
valid <- window(myts, start = c(57, 5)) # October of 2022

# Set Forecast periods
h <- length(valid)


#############
# Train Models
# Trend + Seasonal Model
ST <- forecast::forecast(forecast::tslm(train ~ trend + season),h=h)

# Exponential Model
EXP <- forecast::forecast(forecast::tslm(train ~ trend + season, lambda = 0),h=h)

# NNAR - Autofit neural network
NNAR <- forecast::forecast(nnetar(train), h=h)#, lambda=0)


# Auto ARIMA
ARIMA <- forecast::forecast(auto.arima(train, lambda=0, biasadj=TRUE),h=h)


# Ensemble Approach
Combination <- (EXP[["mean"]] + 
                  ARIMA[["mean"]] +
                  ST[["mean"]] + 
                  NNAR[["mean"]])/4

```


```{r Model Plots}
############
# Baseline Naïve Model

naive_df <- df %>% 
  mutate(naive_forecast = (lag7 + lag14)/2) %>% 
  select(svc_agreement_activation_date, naive_forecast, activation_count)


# Create Time Series Object
mynaivets <- ts(naive_df$naive_forecast, 
                start = c(1,1),
                frequency = 7)

# Plot Results
autoplot(window(myts, start=c(55, 1))) +
  autolayer(window(mynaivets, start=c(57, 5)), series="Baseline Forecast") +
  autolayer(ST, series="Seasonal Trend", PI=FALSE) +
  autolayer(EXP, series="Exponential", PI=FALSE) +
  autolayer(ARIMA, series="ARIMA", PI=FALSE) +
  autolayer(NNAR, series="NNAR", PI=FALSE) +
  autolayer(Combination, series="Combination") +
  xlab("Date") + 
  ylab("Activations") +
  ggtitle("Forecasts for daily activations - 12 Weeks of Training Data") +
  theme_minimal() +
  guides(colour=guide_legend(title="Forecast"))
```

## Including Plots
```{r results}

# Baseline model

resids <- naive_df %>% 
  filter(svc_agreement_activation_date >= as.Date('2022-10-01')) %>% 
  mutate(residuals = naive_forecast - activation_count) %>% 
  select(residuals)

# RMSE
print("Baseline Forecast")
print("RMSE - Test Set")
round(sqrt(mean(resids$residuals^2)),2)
print("")

print(strrep("#", 80))
print("Seasonal Trend")
round(forecast::accuracy(ST, valid),2)
print("")

print(strrep("#", 80))
print("Exponential")
round(forecast::accuracy(EXP, valid),2)
print("")

print(strrep("#", 80))
print("Neural Network")
round(forecast::accuracy(NNAR, valid),2)
print("")    
    
print(strrep("#", 80))
print("ARIMA")
round(forecast::accuracy(ARIMA, valid),2)
print("")

print(strrep("#", 80))
print("Combination")
round(forecast::accuracy(Combination, valid),2)
print("")
```

You can also embed plots, for example:

```{r RMSE}
c(
  Baseline = round(sqrt(mean(resids$residuals^2)),2),
  ST = forecast::accuracy(ST, valid)["Test set","RMSE"],
  EXP = forecast::accuracy(EXP, valid)["Test set","RMSE"],
  NNAR = forecast::accuracy(NNAR, valid)["Test set","RMSE"],
  ARIMA = forecast::accuracy(ARIMA, valid)["Test set","RMSE"],
  Combination = forecast::accuracy(Combination, valid)["Test set","RMSE"])

```

Let's try adding the Baseline Forecast to the ensemble to see if we get an improvement in our prediction.


```{r}


baseline.predict <- df %>% 
  select(svc_agreement_activation_date, activation_count, lag7, lag14) %>% 
  mutate(naive_forecast = (lag7 + lag14)/2) %>% 
  filter(svc_agreement_activation_date >= as.Date('2022-10-01')) 


NewCombination <- (EXP[["mean"]] + 
                  ARIMA[["mean"]] +
                  ST[["mean"]] + 
                  NNAR[["mean"]] + 
                  baseline.predict[["naive_forecast"]])/5

round(forecast::accuracy(NewCombination, valid),2)

```